Distance Sampling in \proglang{R}

### 2016-03-18 paper.Rmd
survey design stuff

The distribution of the observed distances is a product of detectability [sometimes referred to as "perception bias"; @Marsh:1989ho] and the distribution of the animals (observed or unobserved) with respect to the line or point. Our survey design allow us to assume a distribution for the animals with respect to the sampler.

Given that the lines or points are located independently of the objects, for line transect studies we assume that objects are uniformly distributed with respect to the line (i.e., the number of animals available for detection is the same at all distances). For point transect surveys the survey area increases linearly with radial distance, implying a triangular distribution with respect to the point. Figure \ref{fig:detectiondist} shows how these distributions, when combined with a detection function, give rise to the observed distribution of recorded distances.

@@

```{r points-and-lines, fig.height=6, fig.width=7, echo=FALSE, message=FALSE, fig.cap="Left side plots show an example of a survey of an area containing a population of 500 objects; blue dashed lines (top plot) and triangles (bottom plot) indicate sampler placement, red dots indicate detected individuals and grey dots give the locations of unobserved individuals. The right side of the figure shows histograms of observed distances (again, lines top and points bottom).\\label{fig:pointslines}"}
opar <- par(mfrow=c(2,2))
# line and point transect examples
# top row points/samplers in space
# bottom row histograms

@@

par(opar)

>>>>

The distribution of the observed distances is a product of detectability [sometimes referred to as "perception bias"; @Marsh:1989ho] and the distribution of the animals (observed or unobserved) with respect to the line or point. Our survey design allows us to assume a distribution for the animals with respect to the sampler.

For line transect studies we assume that objects are uniformly distributed with respect to the line (i.e., the number of animals available for detection is the same at all distances). For point transect surveys the survey area increases linearly with radial distance, implying a triangular distribution with respect to the point. Figure \ref{fig:detectiondist} shows how these distributions, when combined with a detection function, give rise to the observed distribution of recorded distances.

@@

```{r points-and-lines, fig.height=6, fig.width=7, echo=FALSE, message=FALSE, fig.cap="Left side plots show an example of a survey of a (1x1) area containing a population of 500 objects; blue dashed lines (top plot) and triangles (bottom plot) indicate sampler placement, red dots indicate detected individuals and grey dots give the locations of unobserved individuals. The right side of the figure shows histograms of observed distances (again, lines top and points bottom).\\label{fig:pointslines}"}
opar <- par(mfrow=c(2,2))
# line and point transect examples
# top row points/samplers in space
# bottom row histograms

@@

par(opar)
```

Good survey design is essential to produce reliable density and abundance estimates from distance sampling data. Survey design is beyond the scope of this article but we refer readers to @Buckland:2001vm [Chapter 7] and @buckland2015distance [Chapter 2] for introductory information; @Buckland:2004ts [Chapter 7] contains information on automated survey design using geographical information systems (GIS); @Thomas:2007wz gives an example of designing a distance sampling survey in a complex region.
